Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-01-30 PAPER.tex
some useful preamble bits

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[figuresright]{rotating}
\usepackage{natbib}
\usepackage{ctable}
\usepackage{comment}
%\usepackage{titlesec}
\citestyle{aa}

@@

By about one billion years after the Big Bang ($z \sim 6$), the first stars and galaxies are thought to have ionized all the 
neutral hydrogen that dominated the baryonic matter content in the Universe. This transition period, during which the first 
luminous structures formed from gravitational collapse and began to emit intense radiation that ionized the cold neutral gas 
into a plasma, is known as the Epoch of Reionization (EoR). The EoR is a relatively unexplored era in our \textit{cosmic dawn}, which spans from the birth of the first stars to the full reionization of the intergalactic medium (IGM). Its 
history encodes important information regarding the nature of the first galaxies and the processes of structure formation. 
Direct measurements of the EoR would unlock powerful characteristics about the IGM, revealing connections 
between the matter distribution exhibited via cosmic microwave background (CMB) studies and the highly structured 
web of galaxies we observe today (for a review, see \citet{barkana_and_loeb2001}, \citet{furlanetto_et_al2006} and \citet{loeb_furlanetto_2013}).

One promising technique to probe the EoR is to target the 21\,cm wavelength signal that is emitted and absorbed by neutral hydrogen via 
its spin-flip transition (\citealt{furlanetto_et_al2006}; \citealt{barkana_and_loeb2008}; \citealt{morales_and_wyithe2010}; \citealt{pritchard_and_loeb2010}; \citealt{pritchard_loeb2012}). This tomography technique is powerful because it can be observed both spatially and as a function of redshift --- that is, the wavelength 
of the signal reaching our telescopes can be directly mapped to a distance from where the emission originated before 
stretching out as it traveled through expanding space. Tracing the 21\,cm line as a function of redshift offers a window into the 
evolution of ionization, temperature, and density fluctuations.

@@

The major challenge that faces all 21\,cm experiments is isolating a small signal that is buried underneath foregrounds and 
instrumental systematics that are, when combined, four to five orders of magnitude brighter \citep[e.g.,][]{santos_et_al2005, ali_et_al2008, deOliveiraCosta_et_al2008, jelic_et_al2008, bernardi_et_al2009, bernardi_et_al2010, ghosh_et_al2011, pober_et_al2013, bernardi_et_al2013, dillon_et_al2014, kohn_et_al2016}. A clean measurement therefore requires an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made 
in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact 
with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure 
the accuracy and reliability of a potential EoR detection. In this paper, we discuss three topics (signal loss, errors, and bias) that are essential to investigate 
for a robust $21$ cm power spectrum analysis. We also highlight four power spectrum techniques (fringe-rate filtering, weighting, bootstrapping, jackknife testing) and their trade-offs, potential 
pitfalls, and connections to the themes. We first approach the themes from a broad perspective, and then perform a detailed 
case study using data from the 64-element configuration of PAPER, motivating a revised PAPER-64 power spectrum 
from the lessons learned. More specifically, we develop new techniques for computing signal loss and under-go a rigorous re-analysis of PAPER-64 data in order to present a new power spectrum limit. A companion paper, Kolopanis et al., \textit{in prep}, builds off of the methods in this paper to present PAPER-64 results for multi-redshifts and multiple baselines.

Finally, this paper adds to the growing foundations of lessons which have been documented in  \citet{Patil2016}, \citet{Jacobs2016}, and \cite{Paciga2013} by the LOFAR, MWA and GMRT projects respectively. These lessons are imperative as the community as a whole moves towards higher sensitivities and potential EoR detections.

@@

Driven by the need to mitigate foreground bias, PAPER's previous analyses use a weighting method that aims to down-weight 
foregrounds. This weighting is applied to data, which is then propagated into a final estimator using the power spectrum 
estimation technique of optimal quadratic estimators (OQE) as done in \citet{liu_tegmark2011}, \citet{dillon_et_al2013a}, \citet{liu_et_al2014a}, \citet{liu_et_al2014b}, \citet{trott_et_al2012}, \citet{dillon_et_al2014}, \citet{dillon_et_al2015}, \citet{switzer_et_al2015}, and \citet{trott_et_al2016}. Before showing 
how signal loss can arise when using different weighting matrices, we first summarize OQE as performed in the PAPER 
analysis.

@@

Without a perfect model for the covariance matrix $\textbf{C}$ of our data, one attractive way to estimate it is to empirically derive it from the data $\textbf{x}$ itself. Similar types of weightings that are based on variance information in data are done in \citet{chang_et_al2010} and \citet{switzer_et_al2015}. In previous PAPER analyses, one time-averages the data such that:

@@

The danger of an empirically estimated covariance matrix comes mostly from not being able to describe the `weak' eigenmodes of $
\textbf{C}$ accurately, for which the EoR signal is brighter than foregrounds. In such a case, the coupling between these modes to the data realization leads to the overfitting and subtraction of the EoR signal. More specifically, the coupling between the estimated covariance and the data is anti-correlated in nature (which is explained in more detail in Section \ref{sec:Illustration}), which leads to loss. Mis-estimating $\textbf{C}$ for EoR-dominated eigenmodes is therefore more harmful than for FG-dominated modes, and since the `weak' modes of an eigenspectrum are typically EoR-dominated, using this part of the spectrum for weighting is most dangerous.

@@

Our second major $21$ cm power spectrum theme is error estimation, as we desire robust methods for determining accurate 
confidence intervals for our measurements. Two popular ways of calculating errors on a power spectrum 
measurement are calculating the variance of power spectrum results, and computing a theoretical error estimate based on an instrument's 
system temperature and observational parameters. In a perfect world, both methods would match up. However, in practice the 
two don't always agree due to a number of factors, including possible non-Gaussianities in the noise properties of our instruments and non-uniform 
weightings.

%A third option which acts as a middle ground between the other two is

@@

In Section \ref{sec:SiglossOverview}, we showed how signal loss arises when weighting data using information from the data 
itself. Here we describe a methodology for estimating the amount of signal loss caused by a particular power spectrum estimator when applied to a particular data set. The exact signal loss ``correction factor" for a particular data set will depend on the specific realizations of the signals present in the data and is not something we can directly compute. 
In this work, we inject simulated cosmological signals into our data and test the recovery of those signals in order to statistically quantify 
the amount of signal loss accompanying a weighting scheme used in OQE power spectrum estimation (an approach also taken by \citet{masui_et_al2013}).  
There are, in fact, several different ways of incorporating the result of such a calculation into a revised power spectrum estimate which accounts for signal loss.  In this section, we describe a methodology for determining an approximate correction factor which can be applied to a power spectrum estimate; in Appendix \ref{sec:Pr_appendix}, we present an alternate framework that explicitly interprets all measurements as upper limits in the presence of signal loss.  
These two methods yield very similar results, but fundamentally are answering subtly distinct questions.  
We also highlight major differences from the signal loss 
computation used in A15, which previously underestimated losses.

@@

In the limit of no instrumental noise, the data that we measure, $\x$, is comprised of two signals, such that
\begin{equation}
\x \equiv \f + \s,
\end{equation}
where $\mathbf{f}$ represents the foregrounds and $\mathbf{s}$ represents the cosmological signal. In general, suppose that our power spectrum algorithm passes the data through some function $g$, yielding a lossy estimate of the power spectrum $\phat_{x}$. This can be parametrized as
\begin{equation}
\label{eq:LinearPspecSum}
\langle \phat_{x} \rangle  = \langle g(\x) \rangle = T_{\rm fg} \p_{\rm fg} + T_{\rm eor} \p_{\rm eor},
\end{equation}
where $T_{\rm eor}$ and $T_{\rm fg}$ are multiplicative factors accounting for the signal loss in the true EoR 
power spectrum $\p_{\rm eor}$ and true foreground power spectrum $\p_{\rm fg}$, respectively. It is not 
\emph{a priori} obvious why this parameterization is suitable; we thus provide a toy model in Appendix 
\ref{sec:sigloss_appendix} to motivate this, although it should be noted that the derivation is an approximation 
which assumes that the covariance used in the OQE analysis is close to the true covariance with effects due to 
small sample size limited to first order perturbations and thus neglects cross correlation between $x$ and $e$.  
In actual fact the sample size is similar to the number of independent modes, a case were one expects cross 
terms to be remain significant.

Given this form, a suitable estimate for $\p_{\rm eor}$ would be $\phat_{x} / T_{\rm eor}$, or the uncorrected power spectrum of data divided by the signal loss estimate. Although such an estimate leaves an additive bias from foregrounds (see Section \ref{sec:BiasTypes}) that must be mitigated by other methods (as is the case for any attempt to measure the $21\,\textrm{cm}$ power spectrum), it normalizes $\p_{\rm eor}$ back to its correct level such that there is no multiplicative bias. The most conceptually straightforward way to compute $T_{\rm eor}$ is to model the foregrounds and EoR signal via simulations, and then to form the quantity
\begin{equation}
\label{eq:Deriv1}
\widehat{T}_{\rm eor} = \frac{g(\f + \s) - g(\f)}{\p_{\rm eor}}.
\end{equation}
However, this approach assumes that we have sufficiently good knowledge of our foreground and signal models, which is certainly not the case --- if it were, it would be simpler to construct our covariance matrices from our models, avoiding signal loss altogether! Instead, we can use the data itself as our model of the foregrounds, injecting a new EoR signal $\e$ (with power spectrum $P_{in}$), computing instead
\begin{equation}
\label{eq:Deriv2}
\widehat{T}_{\rm eor} = \frac{g(\x+ \e) - g(\x)}{P_{in}},
\end{equation}
which reduces to the same result because of the linearity of Equation \eqref{eq:LinearPspecSum}. Essentially, one is computing the slope of $\langle \phat_{x} \rangle$ with respect to $\p_{\rm eor}$ (note that both Equations \eqref{eq:Deriv1} and \eqref{eq:Deriv2} take the form of finite difference derivatives). Under the approximation that the relation between the two quantities is linear, it does not matter whether this slope is evaluated about $\x$ or $\f$. In reality, one expects some deviations from linearity, but Equation \eqref{eq:Deriv2} remains a good approximation of Equation \eqref{eq:Deriv1} as long as $\x$ is dominated by $\f$.

@@

The correlation between $\widehat{\textbf{C}}_r^{-1}\textbf{e}$ and $\widehat{\textbf{C}}_r^{-1}\textbf{x}$ means that we cannot compute signal loss using a signal-only 
simulation, which would yield greater values for $P_{\rm out}$ and thereby underestimate signal loss. This argument is also made in \citet{switzer_et_al2015} in regards to estimating loss associated with foreground cleaning of Great Bank Telescope (GBT) data. Therefore, in our revised 
signal loss computation we use the full quantity for $P_{\rm out}$ as defined in Equation \eqref{eq:sigloss}, which subtracts the 
weighted power spectrum of the data from the weighted power spectrum of data plus EoR.

@@

\phat_x \approx T_{\rm fg} |\mathbf{f}|^2 + T_{\rm eor} |\mathbf{e}|^2 + T_{\rm fe} \mathbf{f} \cdot \mathbf{e}.

@@

\phat_x \approx T_{\rm fg} |\mathbf{f}|^2 +( T_{\rm eor} + \rho  T_{\rm fe} ) |\mathbf{e}|^2 + T_{\rm fe} \mathbf{f}\cdot \mathbf{f}_\perp.

@@

\phat_x \approx T_{\rm fg} f^2 + T_{\rm eor} e^2 + T_{\rm fe} fe.

@@

$g(x)$ to estimate $\hat{T}_{\rm eor}$?  How do we subtract off $f\dot f_\perp$? Don't we want to know $
T_{\rm eor}$ not $(T_{\rm eor} + \rho T_{\rm fe})$? }
\acl{I think what we are trying to say here is that actually the thing we want \emph{is} $(T_{\rm eor} + \rho T_{\rm fe})$. Essentially, the potential coupling between the injected EoR and the foregrounds means that we lose the nice form of Equation \eqref{eq:LinearPspecSum}, which justified what is basically a multiplicative correction to signal loss. What we are suggesting here is that even when there is a coupling, the estimator can be parameterized as $\alpha  |\mathbf{e}|^2 + \beta$. So essentially we are defining an effective $T_{\rm eor}$, which is given by $(T_{\rm eor} + \rho T_{\rm fe})$.}

The second issue to tackle is how one incorporates the randomness of $T_{\rm eor}$ into our signal loss corrections. This is manifest in Equation \eqref{eq:fperpfpara}, where the randomness of $\rho$ means that there is an inherent spread in the signal loss correction. Phrased in the context of Bayes' rule, we wish to find the posterior probability distribution of $p( \p_{\rm eor} | \phat_{x})$ for $\p_{\rm eor}$ given the uncorrected power spectrum estimate $\phat_{x}$. Since Equation \eqref{eq:Deriv2} is a good approximation of Equation \eqref{eq:Deriv1}, which we argued for in the previous section, we can write $\p_{\rm eor}$ as $P_{in}$, and hence the probability distribution is given by
\begin{equation}
\label{eq:Bayes}
p(P_{in} | \phat_{x}) \propto \mathcal{L} (  \phat_{x} | P_{in})  p(P_{in}),
\end{equation}
where $p(P_{in})$ is the prior on $P_{in}$ and $\mathcal{L} $ is the likelihood function. Since the likelihood is defined as the distribution of the measured result $ \phat_{x}$ given the theoretical power spectrum $P_{in}$, we may construct this function simply by fixing $P_{in}$, and simulating our analysis pipeline for many realizations of the injected EoR signal consistent with this power spectrum. The resulting distribution can be normalized, and the whole process can then be repeated for a different value of $P_{in}$.

@@

We smooth the 2D distribution of $P_{in}$ vs. $P_{out}$ using kernel density estimators. The result, for inverse covariance weighting, is shown in the left plot of Figure \ref{fig:sigloss_transfercurve}. Bayes' rule then simply instructs us to fix $\phat_{x}$ at our measured value, essentially reading Figure \ref{fig:sigloss_transfercurve} horizontally and reinterpreting Equation \eqref{eq:Bayes} as a function of $p(P_{in})$. The final result can then be normalized to give $p(P_{in} | \phat_{x})$. If $\phat_{x}$ itself does not take on a single value but is itself a distribution (for our analysis, this comes from bootstrapping, where our distribution is shifted so that its peak value aligns with the power spectrum estimate from the no-bootstrapping case; see Section \ref{sec:Boot} for more details on this), then one simply repeats the process for every single point on the distribution of $\phat_{x}$, before performing the final summation and normalization. In Figure \ref{fig:sigloss_transfercurve}, the peak of our data distribution $
\widehat{\textbf{p}}_{x}$ is marked by the solid gray horizontal lines. From the left plot (inverse covariance weighting), one can eyeball that a data value of $10^{5} \, mK^{2} (h^{-1} Mpc)^{3}$, for example, would map approximately to a 
value of $\sim10^{8} \, mK^{2} (h^{-1} Mpc)^{3}$, implying a signal loss factor of $\sim1000$. Performing a summation and normalization for the entire distribution of $\widehat{\textbf{p}}_{x}$ yields a final $P_{\rm in}$ distribution --- the distribution of our data as seen through the signal loss lens. We compute power spectrum points from the peak of the histograms, and power spectrum errors from $95\%$ confidence intervals.

As a final complication to our procedure, we note that there will be some intrinsic scatter in our likelihood due to our having only a finite number of simulations. Ideally, running many simulations would lead to convergence in our likelihood; however, our desire to be able to quickly and easily estimate signal loss for a variety of weighting schemes and datasets means having a finite sample in practice. To separate the intrinsic stochasticity of signal loss from that which arises due to simulation noise, we repeat our analysis for a power spectrum estimator without signal loss ($\textbf{R} = \textbf{I}$), shown as the right plot in Figure \ref{fig:sigloss_transfercurve}. This intrinsic scatter is shown as the smeared `heat-map' around the dotted black diagonal line (representing unity-transfer) in the plot. The width of the lossless scenario's signal loss likelihood is then deconvolved from the lossy scenario's signal loss likelihood. Specifically, we de-convolve Gaussian distributions of $P_{out}$ in the unweighted case from those of the weighted case (vertical cuts through the plots in Figure \ref{fig:sigloss_transfercurve}), for every $P_{in}$. By doing this, we are left with only the intrinsic scatter in signal loss, or scatter that stems from how much the random EoR signal $\textbf{e}$ happens to 
look like the data $\textbf{x}$, a quantity we do not know offhand but one that we would like to correct for. As an extreme example, if we are very unlucky, one realization of $\textbf{e}$ would have the same shapes, or eigenvectors, as $\textbf{x}$. An empirically-derived covariance would then down-weight these shapes, destroying the entire EoR signal. On the other hand, the less that $\textbf{e}$ looks like $\textbf{x}$, the less signal loss that would result. For the case shown in Figure \ref{fig:sigloss_transfercurve}, the intrinsic scatter we obtain from the de-convolution is small or negligible for each $P_{in}$. This means that for our analysis, the intrinsic stochasticity of signal loss is not a dominant factor. However, it is nevertheless important to correct for the fact that a particular $P_{\rm out}$ value could arise from a range of $P_{\rm in}$ values.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_transfercurve_posneg.png}
	\caption{Signal loss transfer functions showing the relationship of $P_{\rm in}$ and $P_{\rm out}$, as defined by Equations \eqref{eq:Pin} and \eqref{eq:sigloss}. Kernel density estimations of the power spectrum transfer functions are shown as colored 
heat-maps for the cases of inverse covariance weighted PAPER-64 data (left) and unweighted data (right). The dotted black 
diagonal lines mark a perfect unity mapping, and the solid gray horizontal line denotes the power spectrum value of the data $\widehat{\textbf{p}}_{x}$. From these plots, it is clear that inverse covariance weighting results in $\sim3$ orders of magnitude of signal loss 
for power spectrum values above $\sim$$2 \times 10^{4}$ $mK^{2} \, (h^{-1} \, Mpc)^{3}$, whereas the unweighted case does 
not exhibit loss. The peculiar `signal gain' tail seen at low injection is driven by the cross-terms in Equation \eqref{eq:pout_expand}, also shown by the red curve in the right panel in Figure \ref{fig:sigloss_terms}.}
	\label{fig:sigloss_transfercurve}
\end{figure*}

@@

We bootstrap PAPER-64 power spectra in order to determine confidence intervals for our results. In this section, we highlight 
one major change in the way we estimate errors since A15, using the lesson we have learned about bootstrapping independent samples.

>>>>

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[figuresright]{rotating}
\usepackage{natbib}
\usepackage{ctable}
\usepackage{comment}
%\usepackage{mathtools}
\usepackage{xparse}
%\usepackage{titlesec}
\citestyle{aa}

%Math stuff from Danny
\NewDocumentCommand{\evalat}{sO{\big}mm}{%
  \IfBooleanTF{#1}
   {\mleft. #3 \mright|_{#4}}
   {#3#2|_{#4}}%
}

@@

By about one billion years after the Big Bang ($z \sim 6$), the first stars and galaxies are thought to have ionized all the 
neutral hydrogen that dominated the baryonic matter content in the Universe. This transition period, during which the first 
luminous structures formed from gravitational collapse and began to emit intense radiation that ionized the cold neutral gas 
into a plasma, is known as the Epoch of Reionization (EoR). The EoR is a relatively unexplored era in our \textit{cosmic dawn}, which spans from the birth of the first stars to the full reionization of the intergalactic medium (IGM). Its 
history encodes important information regarding the nature of the first galaxies and the processes of structure formation. 
Direct measurements of the EoR would unlock powerful characteristics about the IGM, revealing connections 
between the matter distribution exhibited via cosmic microwave background (CMB) studies and the highly structured 
web of galaxies we observe today (for a review, see \citet{furlanetto_et_al2006} and \citet{loeb_furlanetto_2013}).

One promising technique to probe the EoR is to target the 21\,cm wavelength signal that is emitted and absorbed by neutral hydrogen via 
its spin-flip transition (\citealt{furlanetto_et_al2006}; \citealt{morales_and_wyithe2010}; \citealt{pritchard_loeb2012}). This tomography technique is powerful because it can be observed both spatially and as a function of redshift --- that is, the wavelength 
of the signal reaching our telescopes can be directly mapped to a distance from where the emission originated before 
stretching out as it traveled through expanding space. Tracing the 21\,cm line as a function of redshift offers a window into the 
evolution of ionization, temperature, and density fluctuations.

@@

The major challenge that faces all 21\,cm experiments is isolating a small signal that is buried underneath foregrounds and 
instrumental systematics that are, when combined, four to five orders of magnitude brighter \citep[e.g.,][]{santos_et_al2005, ali_et_al2008, deOliveiraCosta_et_al2008, jelic_et_al2008, bernardi_et_al2009, bernardi_et_al2010, ghosh_et_al2011, pober_et_al2013, dillon_et_al2014, kohn_et_al2016}. A clean measurement therefore requires an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made 
in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact 
with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure 
the accuracy and reliability of a potential EoR detection. In this paper, we discuss three topics (signal loss, errors, and bias) that are essential to investigate 
for a robust $21$ cm power spectrum analysis. We also highlight four power spectrum techniques and their trade-offs, potential 
pitfalls, and connections to the themes. We first approach the themes from a broad perspective, and then perform a detailed 
case study using data from the 64-element configuration of PAPER, motivating a revised PAPER-64 power spectrum 
from the lessons learned.

This paper adds to the growing foundations of lessons which have been documented in  \cite{Patil2016,Jacobs2016} and \cite{Paciga2013} by the LOFAR, MWA and GMRT projects respectively.

@@

Driven by the need to mitigate foreground bias, PAPER's previous analyses use a weighting method that aims to down-weight 
foregrounds. This weighting is applied to data, which is then propagated into a final estimator using the power spectrum 
estimation technique of optimal quadratic estimators (OQE) as done in \citet{liu_tegmark2011}, \citet{dillon_et_al2013a}, \citet{liu_et_al2014a}, \citet{liu_et_al2014b}, \citet{trott_et_al2012}, \citet{dillon_et_al2014}, \citet{dillon_et_al2015}, and \citet{trott_et_al2016}. Before showing 
how signal loss can arise when using different weighting matrices, we first summarize OQE as performed in the PAPER 
analysis.

@@

Without a perfect model for the covariance matrix $\textbf{C}$ of our data, one attractive way to estimate it is to empirically derive it from the data $\textbf{x}$ itself (this is what was done in the previous PAPER analyses). For instance, one could time-average such that:

@@

The danger of an empirically estimated covariance matrix comes mostly from not being able to describe the `weak' eigenmodes of $
\textbf{C}$ accurately, for which the EoR signal is brighter than foregrounds. In such a case, the coupling between these modes to the data realization leads to the overfitting and subtraction of the EoR signal. In other words, mis-estimating $\textbf{C}$ for EoR-dominated eigenmodes is more harmful than for FG-dominated modes, and since the `weak' modes of an eigenspectrum are typically EoR-dominated, using this part of the spectrum for weighting is most dangerous.

@@

Our second major $21$ cm power spectrum theme is error estimation, as we desire robust methods for determining accurate 
confidence intervals for our measurements. Two popular ways of calculating errors on a power spectrum 
measurement are calculating the variance of a dataset, and computing a theoretical error estimate based on an instrument's 
system temperature and observational parameters. In a perfect world, both methods would match up. However, in practice the 
two don't always agree due to a number of factors, including possible non-Gaussianities in the noise properties of our instruments and non-uniform 
weightings.

@@

In Section \ref{sec:SiglossOverview}, we showed how signal loss arises when weighting data using information from the data 
itself. Here we describe a methodology for estimating the amount of signal loss caused by a particular power spectrum estimator when applied to a particular data set. The exact signal loss ``correction factor" for a particular data set will depend on the specific realizations of the signals present in the data and is not something we can directly compute. 
In this work, we inject simulated cosmological signals into our data and test the recovery of those signals in order to statistically quantify 
the amount of signal loss accompanying a weighting scheme used in OQE power spectrum estimation.  
There are, in fact, several different ways of incorporating the result of such a calculation into a revised power spectrum estimate which accounts for signal loss.  In this section, we describe a methodology for determining an approximate correction factor which can be applied to a power spectrum estimate; in Appendix \ref{sec:Pr_appendix}, we present an alternate framework that explicitly interprets all measurements as upper limits in the presence of signal loss.  
These two methods yield very similar results, but fundamentally are answering subtly distinct questions.  
We also highlight major differences from the signal loss 
computation used in A15, which previously underestimated losses.

@@

In the limit of no instrumental noise, the data that we measure, $\x$, is comprised of two signals, such that
\begin{equation}
\x \equiv \f + \s,
\end{equation}
where $\mathbf{f}$ represents the foregrounds and $\mathbf{s}$ represents the cosmological signal. In general, suppose that our power spectrum algorithm passes the data through some function $g$, yielding a lossy estimate of the power spectrum $\phat_{x}$. This can be parametrized as
\begin{equation}
\label{eq:LinearPspecSum}
\langle \phat_{x} \rangle  = \langle g(\x) \rangle = \ell_{\rm fg} \p_{\rm fg} + \ell_{\rm eor} \p_{\rm eor},
\end{equation}
where $\ell_{\rm eor}$ and $\ell_{\rm fg}$ are multiplicative factors accounting for the signal loss in the true EoR 
power spectrum $\p_{\rm eor}$ and true foreground power spectrum $\p_{\rm fg}$, respectively. It is not 
\emph{a priori} obvious why this parameterization is suitable; we thus provide a toy model in Appendix 
\ref{sec:sigloss_appendix} to motivate this, although it should be noted that the derivation is an approximation 
which assumes that the covariance used in the OQE analysis is close to the true covariance with effects due to 
small sample size limited to first order perturbations and thus neglects cross correlation between $x$ and $e$.  
In actual fact the sample size is similar to the number of independent modes, a case were one expects cross 
terms to be remain significant.

Given this form, a suitable estimate for $\p_{\rm eor}$ would be $\phat_{x} / \ell_{\rm eor}$, or the uncorrected power spectrum of data divided by the signal loss estimate. Although such an estimate leaves an additive bias from foregrounds (see Section \ref{sec:BiasTypes}) that must be mitigated by other methods (as is the case for any attempt to measure the $21\,\textrm{cm}$ power spectrum), it normalizes $\p_{\rm eor}$ back to its correct level such that there is no multiplicative bias. The most conceptually straightforward way to compute $\ell_{\rm eor}$ is to model the foregrounds and EoR signal via simulations, and then to form the quantity
\begin{equation}
\label{eq:Deriv1}
\widehat{\ell}_{\rm eor} = \frac{g(\f + \s) - g(\f)}{\p_{\rm eor}}.
\end{equation}
However, this approach assumes that we have sufficiently good knowledge of our foreground and signal models, which is certainly not the case --- if it were, it would be simpler to construct our covariance matrices from our models, avoiding signal loss altogether! Instead, we can use the data itself as our model of the foregrounds, injecting a new EoR signal $\e$ (with power spectrum $P_{in}$), computing instead
\begin{equation}
\label{eq:Deriv2}
\widehat{\ell}_{\rm eor} = \frac{g(\x+ \e) - g(\x)}{P_{in}},
\end{equation}
which reduces to the same result because of the linearity of Equation \eqref{eq:LinearPspecSum}. Essentially, one is computing the slope of $\langle \phat_{x} \rangle$ with respect to $\p_{\rm eor}$ (note that both Equations \eqref{eq:Deriv1} and \eqref{eq:Deriv2} take the form of finite difference derivatives). Under the approximation that the relation between the two quantities is linear, it does not matter whether this slope is evaluated about $\x$ or $\f$. In reality, one expects some deviations from linearity, but Equation \eqref{eq:Deriv2} remains a good approximation of Equation \eqref{eq:Deriv1} as long as $\x$ is dominated by $\f$.

@@

The correlation between $\widehat{\textbf{C}}_r^{-1}\textbf{e}$ and $\widehat{\textbf{C}}_r^{-1}\textbf{x}$ means that we cannot compute signal loss using a signal-only 
simulation, which would yield greater values for $P_{\rm out}$ and thereby underestimate signal loss. Therefore, in our revised 
signal loss computation we use the full quantity for $P_{\rm out}$ as defined in Equation \eqref{eq:sigloss}, which subtracts the 
weighted power spectrum of the data from the weighted power spectrum of data plus EoR.

@@

\phat_x \approx \ell_{\rm fg} |\mathbf{f}|^2 + \ell_{\rm eor} |\mathbf{e}|^2 + \ell_{\rm fe} \mathbf{f} \cdot \mathbf{e}.

@@

\phat_x \approx \ell_{\rm fg} |\mathbf{f}|^2 +( \ell_{\rm eor} + \rho  \ell_{\rm fe} ) |\mathbf{e}|^2 + \ell_{\rm fe} \mathbf{f}\cdot \mathbf{f}_\perp.

@@

\phat_x \approx \ell_{\rm fg} f^2 + \ell_{\rm eor} e^2 + \ell_{\rm fe} fe.

@@

$g(x)$ to estimate $\hat{\ell}_{\rm eor}$?  How do we subtract off $f\dot f_\perp$? Don't we want to know $
\ell_{\rm eor}$ not $(\ell_{\rm eor} + \rho\ell_{\rm fe})$? }
\acl{I think what we are trying to say here is that actually the thing we want \emph{is} $(\ell_{\rm eor} + \rho\ell_{\rm fe})$. Essentially, the potential coupling between the injected EoR and the foregrounds means that we lose the nice form of Equation \eqref{eq:LinearPspecSum}, which justified what is basically a multiplicative correction to signal loss. What we are suggesting here is that even when there is a coupling, the estimator can be parameterized as $\alpha  |\mathbf{e}|^2 + \beta$. So essentially we are defining an effective $
\ell_{\rm eor}$, which is given by $(\ell_{\rm eor} + \rho\ell_{\rm fe})$.}

The second issue to tackle is how one incorporates the randomness of $\ell_{\rm eor}$ into our signal loss corrections. This is manifest in Equation \eqref{eq:fperpfpara}, where the randomness of $\rho$ means that there is an inherent spread in the signal loss correction. Phrased in the context of Bayes' rule, we wish to find the posterior probability distribution of $p( \p_{\rm eor} | \phat_{x})$ for $\p_{\rm eor}$ given the uncorrected power spectrum estimate $\phat_{x}$. Since Equation \eqref{eq:Deriv2} is a good approximation of Equation \eqref{eq:Deriv1}, which we argued for in the previous section, we can write $\p_{\rm eor}$ as $P_{in}$, and hence the probability distribution is given by
\begin{equation}
\label{eq:Bayes}
p(P_{in} | \phat_{x}) \propto \mathcal{L} (  \phat_{x} | P_{in})  p(P_{in}),
\end{equation}
where $p(P_{in})$ is the prior on $P_{in}$ and $\mathcal{L} $ is the likelihood function. Since the likelihood is defined as the distribution of the measured result $ \phat_{x}$ given the theoretical power spectrum $P_{in}$, we may construct this function simply by fixing $P_{in}$, and simulating our analysis pipeline for many realizations of the injected EoR signal consistent with this power spectrum. The resulting distribution can be normalized, and the whole process can then be repeated for a different value of $P_{in}$.

@@

We smooth the 2D distribution of $P_{in}$ vs. $P_{out}$ using kernel density estimators. The result, for inverse covariance weighting, is shown in the left plot of Figure \ref{fig:sigloss_transfercurve}. Bayes' rule then simply instructs us to fix $\phat_{x}$ at our measured value, essentially reading Figure \ref{fig:sigloss_transfercurve} horizontally and reinterpreting Equation \eqref{eq:Bayes} as a function of $p(P_{in})$. The final result can then be normalized to give $p(P_{in} | \phat_{x})$. If $\phat_{x}$ itself does not take on a single value but is itself a distribution (for our analysis, this comes from bootstrapping), then one simply repeats the process for every single point on the distribution of $\phat_{x}$, before performing the final summation and normalization. In Figure \ref{fig:sigloss_transfercurve}, the peak of our data distribution $
\widehat{\textbf{p}}_{x}$ is marked by the solid gray horizontal lines. From the left plot (inverse covariance weighting), one can eyeball that a data value of $10^{5} \, mK^{2} (h^{-1} Mpc)^{3}$, for example, would map approximately to a 
value of $\sim10^{8} \, mK^{2} (h^{-1} Mpc)^{3}$, implying a signal loss factor of $\sim1000$. Performing a summation and normalization for the entire distribution of $\widehat{\textbf{p}}_{x}$ yields a final $P_{\rm in}$ distribution --- the distribution of our data as seen through the signal loss lens. We compute power spectrum points from the peak of the histograms, and power spectrum errors from $95\%$ confidence intervals.

As a final complication to our procedure, we note that there will be some intrinsic scatter in our likelihood due to our having only a finite number of simulations. Ideally, running many simulations would lead to convergence in our likelihood; however, our desire to be able to quickly and easily estimate signal loss for a variety of weighting schemes and datasets means having a finite sample in practice. To separate the intrinsic stochasticity of signal loss from that which arises due to simulation noise, we repeat our analysis for a power spectrum estimator without signal loss ($\textbf{R} = \textbf{I}$), shown as the right plot in Figure \ref{fig:sigloss_transfercurve}. This intrinsic scatter is shown as the smeared `heat-map' around the solid black diagonal line (representing unity-transfer) in the plot. The width of the lossless scenario's signal loss likelihood is then deconvolved from the lossy scenario's signal loss likelihood. Specifically, we deconvolve Gaussian distributions of $P_{out}$ in the unweighted case from those of the weighted case (vertical cuts through the plots in Figure \ref{fig:sigloss_transfercurve}), for every $P_{in}$. By doing this, we are left with only the intrinsic scatter in signal loss, or scatter that stems from how much the random EoR signal $\textbf{e}$ happens to 
look like the data $\textbf{x}$, a quantity we do not know offhand but one that we would like to correct for. As an extreme example, if we are very unlucky, one realization of $\textbf{e}$ would have the same shapes, or eigenvectors, as $\textbf{x}$. An empirically-derived covariance would then down-weight these shapes, destroying the entire EoR signal. On the other hand, the less that $\textbf{e}$ looks like $\textbf{x}$, the less signal loss that would result. The intrinsic scatter we can get is not a dominant factor in this case but it is important to correct for the fact that a particular $P_{\rm out}$ value could arise from a range of $P_{\rm in}$ values.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_transfercurve.png}
	\caption{Signal loss transfer functions showing the relationship of $P_{\rm in}$ and $P_{\rm out}$, as defined by Equations \eqref{eq:Pin} and \eqref{eq:sigloss}. Kernel density estimations of the power spectrum transfer functions are shown as colored 
heat-maps for the cases of inverse covariance weighted PAPER-64 data (left) and unweighted data (right). The solid black 
diagonal line marks a perfect unity mapping, and the solid gray horizontal line denotes the power spectrum value of the data $\widehat{\textbf{p}}_{x}$. From these plots, it is clear that inverse covariance weighting results in $\sim3$ orders of magnitude of signal loss 
for power spectrum values above $\sim$$2 \times 10^{4}$ $mK^{2} \, (h^{-1} \, Mpc)^{3}$, whereas the unweighted case does 
not exhibit loss. The peculiar `signal gain' tail seen at low injection is driven by the cross-terms in Equation \eqref{eq:pout_expand}, also shown by the red curve in the right panel in Figure \ref{fig:sigloss_terms}.}
	\label{fig:sigloss_transfercurve}
\end{figure*}

@@

We bootstrap PAPER-64 power spectra in order to determine confidence intervals for our results. In this section, we highlight 
one major changes in the way we estimate errors since A15, using the lesson we have learned about bootstrapping independent samples.
